---
title: "Reproducibility Review AGILE 2020"
author: "Daniel Nüst"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    self_contained: true
  #pdf_document:
  #  toc: yes
params:
  private_info: yes
---

This document includes some scripts and text analysis to support the reproducibility review at the [AGILE conference 2020](https://agile-online.org/conference-2020), in Chania, Greece - _"Geospatial Technologies: seeding the future"_.

Find out more about the reproducible publications at AGILE [here](https://doi.org/10.17605/OSF.IO/PHMCE) all about the review process at https://osf.io/eg4qx/.
You can see the source code of the analysis in the R Markdown file [`agile-2020-papers.Rmd`](agile-2020-papers.Rmd).

```{r load_libraries, message=FALSE, warning=FALSE, include=FALSE}
library("pdftools")
library("stringr")
library("tidyverse")
library("tidytext")
library("wordcloud")
library("RColorBrewer")
library("here")
library("quanteda")
library("googledrive")
library("kableExtra")
```

```{r seed, echo=FALSE}
set.seed(23) # 23rd AGILE!
```

## Submitted full papers

```{r fp_path, echo=FALSE}
fp_path <- here::here("2020", "full_papers")
#dir.create(fp_path, recursive = TRUE)
```

### Download files

_The submitted full papers cannot be publicly shared._
If you have access, put them in the directory `r fp_path`.

```{r download_fp, echo=FALSE, eval=FALSE}
drive_auth()
fp_drive_dir <- drive_get("https://drive.google.com/drive/folders/1KcSjirs7pHXE88DfKK4LhEaGK4u4H-j_")
fp_drive_dir_contents <- drive_ls(fp_drive_dir)
for (i in rownames(fp_drive_dir_contents)) {
  current_fp_drive <- fp_drive_dir_contents[i,]
  drive_download(as_id(current_fp_drive$id), file.path(fp_path, current_fp_drive$name))
}
```

### Load texts

```{r load_files, echo=FALSE}
fp_files <- dir(path = fp_path, pattern = ".pdf$", full.names = TRUE)

id_from_path <- function(pre, path) {
  name <- basename(path)
  name <- str_replace(name, "AGILE2020_paper_", "") # remove common part of filename
  name <- sapply(X = name, FUN = str_replace_all, pattern = "[_.-]|pdf", replacement = "")
  name <- str_pad(name, width = 3, side = "left", pad = "0")
  paste0(pre, "_", name)
}

fp_data <- tibble(id = id_from_path(pre = "fp", path = fp_files),
                  path = fp_files,
                  type = "full")
#all_files <- bind_rows(fp_data, sp_data, po_data) %>%
#  mutate(type = as.factor(type))
all_files <- fp_data
```

The text is extracted from PDFs and it is processed to create a [tidy](https://www.jstatsoft.org/article/view/v059i10) data structure without [stop words](https://en.wikipedia.org/wiki/Stop_words).
The stop words include specific words, which might be included in the page header, abbreviations, and terms particular to scientific articles, such as `figure`.

```{r tidy_data, echo=FALSE}
texts <- lapply(all_files$path, pdf_text)
texts <- unlist(lapply(texts, str_c, collapse = TRUE))
infos <- lapply(all_files$path, pdf_info)

tidy_texts <- tibble(id = all_files$id,
                     path = all_files$path,
                     type = all_files$type,
                     text = texts,
                     pages = map_chr(infos, function(info) {info$pages}))

# create a table of all words
all_words <- tidy_texts %>%
  select(id,
         type,
         text) %>%
  unnest_tokens(word, text)

# remove stop words and remove numbers
my_stop_words <- tibble(
  word = c(
    "et",
    "al",
    "fig",
    "e.g",
    "i.e",
    "http",
    "ing",
    "pp",
    "figure",
    "based",
    "university",
    "table"
  ),
  lexicon = "agile"
)

all_stop_words <- stop_words %>%
  bind_rows(my_stop_words)
suppressWarnings({
  no_numbers <- all_words %>%
    filter(is.na(as.numeric(word)))
})

no_stop_words <- no_numbers %>%
  anti_join(all_stop_words, by = "word")

total_words = nrow(all_words)
after_cleanup = nrow(no_stop_words)
```

About `r round(after_cleanup/total_words * 100)`&nbsp;% of the words are considered stop words.

The following tables shows how many words and non-stop words each document has, sorted by number of non-stop words.
The `id` is built from the file name plus a prefix:
for full papers, it is the left-padded submission number and the prefix `fp_`;
<!--for short papers and posters, it is the submission number included in the file name and the prefixes `sp_` and `po_` respectively.-->

```{r stop_words, echo=FALSE, message=FALSE, warning=FALSE}
nsw_per_doc <- no_stop_words %>%
  group_by(id) %>%
  summarise(words = n()) %>%
  rename(`non-stop words` = words)

words_per_doc <- all_words %>%
  group_by(id) %>%
  summarise(words = n())

inner_join(words_per_doc, nsw_per_doc) %>%
  select(-starts_with("type")) %>%
  bind_rows(tibble(id = "Total", words = sum(words_per_doc$words), `non-stop words` = sum(nsw_per_doc$`non-stop words`))) %>%
  kable() %>%
  kable_styling("striped", full_width = FALSE) %>%
  row_spec(nrow(nsw_per_doc) + 1, bold = TRUE) %>%
  scroll_box(height = "240px")
```

### Which papers include a "Data and Software Availability" section?

According the the [AGILE Reproducible Paper Guidelines](https://osf.io/c8gtq/), all authors must add a _Data and Software Availability_ section to their paper.
The guidelines are not mandatory yet in 2020, but let's see how many authors did include the statement.

```{r pdfgrep, echo=FALSE, eval=FALSE}
# Quick version with `pdfgrep`
cmd <- paste("pdfgrep", "-e 'Data and Software Availability'", "-A 3", "2020/full_papers/*")
output <- system(cmd, intern = TRUE)
print(cmd)
print(output)
```

```{r dasa_section, echo=FALSE}
tidy_texts <- tidy_texts %>%
  mutate(has_dasa = str_detect(tidy_texts$text, pattern = "Data and Software Availability"))

dasa_count <- tidy_texts %>% filter(has_dasa) %>% nrow()

excerpt_length <- 800
dasa_texts <- tidy_texts %>%
  filter(has_dasa) %>%
  mutate(dasa_start = str_locate(.data$text, pattern = "Data and Software Availability")[,1]) %>%
  mutate(dasa_text = str_sub(.data$text, start = dasa_start, end = dasa_start + excerpt_length)) %>%
  select(id, dasa_text)
```

`r dasa_count` papers have the section in question, that is `r round(dasa_count/nrow(all_files) * 100)`&nbsp;% of all submissions.
The first `r excerpt_length` characters of these sections are as follows.

```{r dasa_section_table, echo=FALSE}
dasa_texts %>%
  kable() %>%
  kable_styling("striped") %>%
  scroll_box(height = "320px")
```


### Wordstem analysis

```{r wordstem_data, include=FALSE}
wordstems <- no_stop_words %>%
  mutate(wordstem = quanteda::char_wordstem(no_stop_words$word))

countPapersUsingWordstem <- function(the_word) {
  sapply(the_word, function(w) {
    wordstems %>%
      filter(wordstem == w) %>%
      group_by(id) %>%
      count %>%
      nrow
  })
}

top_wordstems <- wordstems %>%
  group_by(wordstem) %>%
  tally %>%
  arrange(desc(n)) %>%
  head(20) %>%
  mutate(`# papers` = countPapersUsingWordstem(wordstem)) %>%
  mutate(`% papers` = round(countPapersUsingWordstem(wordstem)/nrow(all_files) * 100)) %>%
  add_column(place = c(1:nrow(.)), .before = 0)

minimum_occurence <- 100
cloud_wordstems <- wordstems %>%
  group_by(wordstem) %>%
  tally %>%
  filter(n >= minimum_occurence) %>%
  arrange(desc(n))
```

For the following table and figure, the word stems were extracted based on a stemming algorithm from package [`quanteda`](https://cran.r-project.org/package=quanteda).
The word cloud is based on `r length(unique(cloud_wordstems$wordstem))` unique words occuring each at least `r minimum_occurence` times, all in all occuring `r sum(cloud_wordstems$n)` times which comprises `r round(sum(cloud_wordstems$n)/ nrow(no_stop_words) * 100)`&nbsp;% of non-stop words.

```{r top_wordstems, echo=FALSE}
top_wordstems %>%
  kable() %>%
  kable_styling("striped") %>%
  scroll_box(height = "320px")
```

```{r wordstemcloud, dpi=150, echo=FALSE, fig.cap="Wordstem cloud of AGILE 2020 full paper submissions"}
wordcloud(cloud_wordstems$wordstem, cloud_wordstems$n,
          max.words = Inf,
          random.order = FALSE,
          fixed.asp = FALSE,
          rot.per = 0,
          color = brewer.pal(8,"Dark2"))
```

## Reproducible research-related keywords

The following tables lists how often terms related to reproducible research appear in each document.
The detection matches full words using regex option `\b`.

- reproduc (`reproduc.*`, reproducibility, reproducible, reproduce, reproduction)
- replic (`replicat.*`, i.e. replication, replicate)
- repeatab (`repeatab.*`, i.e. repeatability, repeatable)
- software
- (pseudo) code/script(s) [column name _code_]
- algorithm (`algorithm.*`, i.e. algorithms, algorithmic)
- process (`process.*`, i.e. processing, processes, preprocessing)
- data (`data.*`, i.e. dataset(s), database(s))
- result(s) (`results?`)
- repository(ies) (`repositor(y|ies)`)

```{r keywords_per_paper, echo=FALSE, warning=FALSE}
tidy_texts_lower <- str_to_lower(tidy_texts$text)
word_counts <- tibble(
  id = tidy_texts$id,
  type = tidy_texts$type,
  DASA = tidy_texts$has_dasa,
  `reproduc..` = str_count(tidy_texts_lower, "\\breproduc.*\\b"),
  `replic..` = str_count(tidy_texts_lower, "\\breplicat.*\\b"),
  `repeatab..` = str_count(tidy_texts_lower, "\\brepeatab.*\\b"),
  `code` = str_count(tidy_texts_lower,
    "(\\bcode\\b|\\bscript.*\\b|\\bpseudo\ code\\b)"),
  software = str_count(tidy_texts_lower, "\\bsoftware\\b"),
  `algorithm(s)` = str_count(tidy_texts_lower, "\\balgorithm.*\\b"),
  `(pre)process..` = str_count(tidy_texts_lower, 
                "(\\bprocess.*\\b|\\bpreprocess.*\\b|\\bpre-process.*\\b)"),
  `data.*` = str_count(tidy_texts_lower, "\\bdata.*\\b"),
  `result(s)` = str_count(tidy_texts_lower, "\\bresults?\\b"),
  `repository/ies` = str_count(tidy_texts_lower, "\\brepositor(y|ies)\\b")
)

# https://stackoverflow.com/a/32827260/261210
sumColsInARow <- function(df, list_of_cols, new_col) {
  df %>% 
    mutate_(.dots = ~Reduce(`+`, .[list_of_cols])) %>% 
    setNames(c(names(df), new_col))
}

word_counts_sums <- sumColsInARow(
  word_counts, 
  names(word_counts)[!(names(word_counts) %in% c("id", "type"))], "all") %>%
  arrange(desc(all))

word_counts_sums_total <- word_counts_sums %>% 
  summarise_if(is.numeric, funs(sum)) %>%
  add_column(id = "Total", type = "", DASA = "", .before = 0)
word_counts_sums <- rbind(word_counts_sums, word_counts_sums_total)

word_counts_sums %>%
  kable() %>%
  kable_styling("striped", font_size = 12, bootstrap_options = "condensed")  %>%
  row_spec(0, font_size = "x-small", bold = T)  %>%
  row_spec(word_counts_sums %>% rownames_to_column() %>%
             filter(DASA == TRUE, .preserve = TRUE) %>%
             select(rowname) %>% unlist() %>% as.numeric(),
           italic = TRUE, background = "#eeeeee") %>%
  row_spec(nrow(word_counts_sums), bold = T)
```

## License & Metadata

This document is licensed under a [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).
All contained code is licensed under the [Apache License 2.0](https://choosealicense.com/licenses/apache-2.0/).

**Runtime environment description:**

```{r session_info}
devtools::session_info(include_base = TRUE)
```

```{r upload_to_drive, eval=FALSE, include=FALSE}
# upload the HTML file to the Reproducibility Committee shared folder
drive_put("agile-2020-papers.html", path = "https://drive.google.com/drive/folders/1jyYj1hFqbR74D9ljjjcScR4lD3aWCO2U/")
```
